Biases in Planet Occurrence Caused by Unresolved Binaries in Transit Surveys

### 2018-04-17 paper/main.tex
add KM's comments

Here, though, we will focus exclusively on problems that arise from
the fact that many stars exist in unresolved binary systems.  We will
also focus on the errors in planet occurrence as a function of radius,
rather than orbital period. This is because when more than one transit
is detected (as is usually required by the surveyors), the orbital
periods can be measured without ambiguity regardless of whether the
host star is single or one member of a binary.  Even with this narrow
focus, there are numerous sources of error.  All three of the
quantities in Equation~\ref{eq:occ_rate_simple} are biased:
\begin{enumerate}
%    
    \item The number of detected planets, $N_{\det}$, is actually the
      number of detected planets that {\it appear} to have size $r$,
      orbiting stars that {\it appear} to have mass $M$.  Whenever the
      planet-hosting star is part of a binary,
%    
    \begin{itemize}
        \item the planet's size could be misclassified because of the
          reduction in the amplitude of the photometric signal;
%        
        \item the host star's properties could be misclassified
          because its light is combined with a second star of a
          different spectral type.
%        
    \end{itemize}
%    
    \item The number of searchable stars, $N_\star$, is biased
%    
    \begin{itemize}
%        
        \item toward lower values, because it does not include all of
          the secondary stars that were inadvertently searched for
          transiting planets;
%        
        \item toward higher values, because some of the stars that
          appeared to be searchable are in fact binaries for which the
          amplitude of the photometric signal would have been reduced
          to an undetectable level.
%        
    \end{itemize}
%    
    \item The transit probability $p_{\rm tra}$ is biased because the
      planet-hosting star could be misclassified.  At fixed orbital
      period, the transit probability is proportional to
      $\rho^{-1/3}$, where $\rho$ is the stellar mean
      density~\citep{Winn2010}.  Therefore, any 
      \deleted{errors}\added{uncertainties} in determining
      the host star's mean density lead to 
      \deleted{errors}\added{uncertainties} in the correction
      for the transit probability.
%    
\end{enumerate}

@@

To compute the apparent occurrence rate density, we need to make
the following modifications to Equation~\ref{eq:formula_twin_binary}:
%
\begin{enumerate}
%
\item The Malmquist bias is different.  For a given value of $r\a$, the
  number of binary systems in the searchable sample with mass ratio
  $(q, q+\mathrm{d}q)$ is
\begin{equation}
    {N_\star(r\a)\over{1+\mu}}\, {n_{\rm b}\over n_{\rm s}}
    \left[{L(M) + L(qM) \over L(M)}\right]^{3/2}
    f(q)\,\mathrm{d}q,
    \label{eq:N_binaries}
\end{equation}
where $\mu$ is given by
\begin{equation}
    \mu= 
    \int_0^1 {n_{\rm b}\over n_{\rm s}}\left[{L(M) + L(qM)
    \over L(M)}\right]^{3/2} f(q)\,\mathrm{d}q.
    \label{eq:mu_general}
\end{equation}
Equation~\ref{eq:N_binaries} replaces ${\mu N_\star(r\a)/(1+\mu)}$
in Equations~\ref{eq:n1} and~\ref{eq:n2}.
The whole equation is then integrated to give the number of
detected planets.

@@

\paragraph{Radius corrections} Another application of our formalism is
to help in those cases in which it is not clear whether the primary or
secondary star is the source of a transit signal.  By this point
almost all of the {\it Kepler}\ Objects of Interest have been examined
with high resolution imaging~\citep{
  howell_speckle_2011,adams_adaptive_2012,adams_adaptive_2013,horch_observations_2012,
  horch_most_2014,lillo-box_multiplicity_2012,lillo-box_high-resolution_2014,dressing_adaptive_2014,
  law_robotic_2014,cartier_revision_2015,everett_high-resolution_2015,gilliland_hubble_2015,
  wang_influence_2015,wang_influence_2015-1,baranec_robo-ao_2016,ziegler_robo-ao_2017}.
The results of these programs have been summarized
by~\citet{furlan_kepler_2017}. Based on this information,
\citet{hirsch_assessing_2017} re-assessed the radius measurements of
the {\it Kepler} planets in binaries. They found that planet radii are
generally underestimated by a factor $r/r\a = 1.17$ if all planets
orbit primaries, and $r/r\a = 1.65$ if detected planets are equally
likely to orbit primaries and singles.  Our formalism provides a way
to assess the relative probabilities that a detected planet orbits the
primary or secondary star.  The calculations presented in
\S\,\ref{subsec:whichstar} suggest that in almost all cases the planet
most likely orbits the primary star.

\added{\paragraph{Caveats \& outlook}
Unresolved binaries corrupt occurrence rate distributions along
dimensions other than planet radius. 
In particular, since binarity can lead to incorrect
assessment of a planet's host stellar mass, the occurrence
distributions in semi-major axis and insolation must also be corrected.
Though we leave consideration of such biases for future work,
we take solace in the fact that the NASA {\it TESS} mission
emphasizes stars within a few hundred parsecs.
Through a combination of high-resolution imaging and 
{\it Gaia}'s astrometry and spectroscopy, it will be possible
to characterize the {\it TESS} target stars more thoroughly than
was possible with {\it Kepler}.
This is essential for inferring properties of planet populations, since
a population inference model with parameters controlling both the planet 
and binary star populations might be necessary to consistently 
account for binarity.
Given the difficulty of writing a correct likelihood function for such
a model, one plausible path might be to
forward-model transit surveys with likelihood-free 
techniques~\citep[\textit{e.g.},][]{morehead_understanding_2016,hsu_improving_2018}.
}

>>>>

Here, though, we will focus exclusively on problems that arise from
the fact that many stars exist in unresolved binary systems.  We will
also focus on the errors in planet occurrence as a function of radius,
rather than orbital period. This is because when more than one transit
is detected (as is usually required by the surveyors), the orbital
periods can be measured without ambiguity regardless of whether the
host star is single or one member of a binary.  Even with this narrow
focus, there are numerous sources of error.  All three of the
quantities in Equation~\ref{eq:occ_rate_simple} are biased:
\begin{enumerate}
%    
    \item The number of detected planets, $N_{\det}$, is actually the
      number of detected planets that {\it appear} to have size $r$,
      orbiting stars that {\it appear} to have mass $M$ \added{and
      radius $R$}.  Whenever the planet-hosting star is part of a
      binary,
%    
    \begin{itemize}
        \item the planet's size could be misclassified because of the
          reduction in the amplitude of the photometric signal;
%        
        \item the host star's properties could be misclassified
          because its light is combined with a second star of a
          different spectral type.
%        
    \end{itemize}
%    
    \item The number of searchable stars, $N_\star$, is biased
%    
    \begin{itemize}
%        
        \item toward lower values, because it does not include all of
          the secondary stars that were inadvertently searched for
          transiting planets;
%        
        \item toward higher values, because some of the stars that
          appeared to be searchable are in fact binaries for which the
          amplitude of the photometric signal would have been reduced
          to an undetectable level.
%        
    \end{itemize}
%    
    \item The transit probability $p_{\rm tra}$ is biased because the
      planet-hosting star could be misclassified.  At fixed orbital
      period, the transit probability is proportional to
      $\rho^{-1/3}$, where $\rho$ is the stellar mean
      density~\citep{Winn2010}.  Therefore, any 
      \deleted{errors}\added{uncertainties} in determining
      the host star's mean density lead to 
      \deleted{errors}\added{uncertainties} in the correction
      for the transit probability.
%    
\end{enumerate}

@@

To compute the apparent occurrence rate density, we need to make
the following modifications to Equation~\ref{eq:formula_twin_binary}:
%
\begin{enumerate}
%
\item The Malmquist bias is different.  For a given value of $r\a$, the
  number of binary systems in the searchable sample with mass ratio
  $(q, q+\mathrm{d}q)$ is
\begin{equation}
    {N_\star(r\a)\over{1+\mu}}\, {n_{\rm b}\over n_{\rm s}}
    \left[{L(M) + L(qM) \over L(M)}\right]^{3/2}
    f(q)\,\mathrm{d}q,
    \label{eq:N_binaries}
\end{equation}
where $\mu$ is given by
\begin{equation}
    \mu= 
    \int_0^1 {n_{\rm b}\over n_{\rm s}}\left[{L(M) + L(qM)
    \over L(M)}\right]^{3/2} f(q)\,\mathrm{d}q.
    \label{eq:mu_general}
\end{equation}
Equation~\ref{eq:N_binaries} replaces ${\mu N_\star(r\a)/(1+\mu)}$
in Equations~\ref{eq:n1} and~\ref{eq:n2}.
The whole equation is then integrated \added{over $q$} to give the
number of detected planets.

@@

\paragraph{Radius corrections} Another application of our formalism is
to help in those cases in which it is not clear whether the primary or
secondary star is the source of a transit signal.  By this point
almost all of the {\it Kepler}\ Objects of Interest have been examined
with high resolution imaging~\citep{
  howell_speckle_2011,adams_adaptive_2012,adams_adaptive_2013,horch_observations_2012,
  horch_most_2014,lillo-box_multiplicity_2012,lillo-box_high-resolution_2014,dressing_adaptive_2014,
  law_robotic_2014,cartier_revision_2015,everett_high-resolution_2015,gilliland_hubble_2015,
  wang_influence_2015,wang_influence_2015-1,baranec_robo-ao_2016,ziegler_robo-ao_2017}.
The results of these programs have been summarized
by~\citet{furlan_kepler_2017}. Based on this information,
\citet{hirsch_assessing_2017} re-assessed the radius measurements of
the {\it Kepler} planets in binaries. They found that planet radii are
generally underestimated by a factor $r/r\a = 1.17$ if all planets
orbit primaries, and $r/r\a = 1.65$ if detected planets are equally
likely to orbit primaries and \deleted{singles}\added{secondaries}.
Our formalism provides a way to assess the relative probabilities that
a detected planet orbits the primary or secondary star.  The
calculations presented in \S\,\ref{subsec:whichstar} suggest that in
almost all cases the planet most likely orbits the primary star.

\added{\paragraph{Caveats \& outlook}
Unresolved binaries corrupt occurrence rate distributions along
dimensions other than planet radius. 
In particular, since binarity can lead to incorrect
assessment of a planet's host stellar mass, the occurrence
distributions in semi-major axis and insolation must also be corrected.
We leave consideration of such biases for future work.}

\added{
It seems likely that to consistently account for binarity in
planet occurrence rate calculation, one must stipulate a
population inference model with parameters controlling both the
planet and binary star populations.
Given the difficulty of writing a correct likelihood function
for such a model, one feasible computational approach might be to
forward-model transit surveys with likelihood-free 
techniques~\citep[\textit{e.g.},][]{morehead_understanding_2016,hsu_improving_2018}.
The NASA {\it TESS} mission should provide ample opportunity
for such a study.
}
